---
title: "brad_la_data_analysis"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
library(foreach)
library(iterators)
library(rjags)
```
## Problems so far:

* DfE covid data: reporting not mandatory - lots of missing schools - bias towards missing data in treatment arm
* Brad Council data v1: Identified errors in possible sessions calculations - issue with aggregation script - OBVIOUSLY NO QA!!
* Brad Council data v2: possible sessions stat net of covid absences plus other authorised absences - can't work out proportion of sessions missed due to covid vs possible others using this data
* Brad Council full data: TBC
* None of the schools seem to follow the covid codings to the letter e.g. forced to self isolate specifically says only 18+ y/o students should fall into this category

```{r}
# load attendance data
attendance <- read_csv("/Users/samrelins/Documents/BIHR/dfe_covid/data/pre_individual_data/brad_la_school_attendance_full.csv")

# add study arms to data
attendance$arm <- 0
arm1 = c(
  "Horton Grange Primary School", "Bradford Academy", 
  "Cavendish Primary School", "St Matthew's Catholic Primary School", 
  "Farnham Primary School", "St Joseph's Catholic Primary School (Bingley)", 
  "Green Lane Primary School", "Westbourne Primary School", 
  "Margaret McMillan Primary School", "Beckfoot Nessfield", 
  "Beckfoot Priestthorpe Primary School & Nursery"
)
attendance[attendance$school %in% arm1, "arm"] <- 1
arm2 = c(
  "Eastwood Community School", "Shirley Manor Primary School",  
  "Swain House Primary School", "Bowling Park Primary School", 
  "Burley and Woodhead CofE Primary School", "Crossflatts Primary School", 
  "Copthorne Primary School", "St Columba's Catholic Primary School", 
  "St Joseph's Catholic Primary School (Bradford)", 
  "St William's Catholic Primary School"
)
attendance[attendance$school %in% arm2, "arm"] <- 2
arm3 = c(
  "Riddlesden St Mary's CofE Primary School", "Newby Primary School", 
  "Thackley Primary School", "Addingham Primary School", 
  "The Academy At St. James", "St Paul's CofE Primary School", 
  "Clayton Village Primary School", "Sandal Primary School", 
  "Grove House Primary School", "Saltaire Primary School"
)
attendance[attendance$school %in% arm3, "arm"] <- 3
attendance <- attendance %>% mutate(hepa_arm = arm == 2)

# convert week_start to date format
attendance <- attendance %>% 
  mutate(week_start = as.Date(week_start, format="%m/%d/%y"))

# remove stray october date
attendance <- attendance %>% filter(week_start > "2021-09-01")

# rename St Joseph's schools to make easier to differentiate
attendance <- attendance %>% 
  mutate(school = replace(school, 
                          school == "St Joseph's Catholic Primary School (Bingley)",  
                          "St Joseph's Bingley")) %>%
  mutate(school = replace(school, 
                          school == "St Joseph's Catholic Primary School (Bradford)",  
                          "St Joseph's Bradford"))

# shorten down school names
shorten_name <- function(name) {
    words = unlist(strsplit(name, " "))
    if(length(words) >= 3) {
      paste(words[1:3], collapse=" ")
    } else {
      paste(words, collapse=" ")
    }
}
attendance <- attendance %>% 
  mutate(school = sapply(school, shorten_name))

# add extra aggregate data and reorder
attendance <- attendance %>%
  mutate(other_attendance = rowSums(across(c(L, B, D, J, P, V, W)))) %>%
  mutate(covid_absences = rowSums(across(c(I02, X06, X07, X02, X09, X08, X)))) %>%
  mutate(illness_absences = rowSums(across(c(I, I01, I02)))) %>%
  mutate(cov_illness_absences = rowSums(across(c(I, I01, covid_absences)))) %>%
  mutate(recorded_sessions = possible_sessions + covid_absences + I01 + Y + `#`) %>%
  mutate(missing_sessions = total_sessions - recorded_sessions) %>%
  relocate(other_attendance, .after=pm) %>%
  relocate(covid_absences, .after=U) %>%
  relocate(recorded_sessions, .after=total_sessions) %>%
  relocate(missing_sessions, .after=recorded_sessions)

str(attendance)
```
## Sanity Checks

```{r}
attendance %>% group_by(week_start) %>% count()
```

Authorised = sum of rows C,E,H,I,M,R,S,T
```{r}
all(unlist(
  attendance %>%
  mutate(check = authorised == rowSums(across(c(C,E,H,I,M,R,S,T)))) %>%
  select(check)
))
```

Unauthorised = sum of rows G,N,O,U
```{r}
all(unlist(
  attendance %>%
  mutate(check = unauthorised == rowSums(across(c(G,N,O,U)))) %>%
  select(check)
))
```

sessions_attended = am + pm + other_attendance
```{r}
all(unlist(
  attendance %>%
  mutate(check = sessions_attended == am + pm + other_attendance) %>%
  select(check)
))
```

possible_sessions = sessions_attended + authorised + unauthorised
```{r}
all(unlist(
  attendance %>%
  mutate(check = possible_sessions == sessions_attended + authorised + unauthorised) %>%
  select(check)
))
```
## Total Sessions Checks

```{r}
attendance %>% filter(school == "Bradford Academy") %>% select(I, I01)
```

```{r}
attendance %>% filter(I01 > 50) %>%
  select(school, week_start, missing_sessions, I, I01)
```

## Main issue - what figure should we use to calculate the proportion of covid absences

* We're evaluating measures to improve air quality in shcools - so, need a proportion of children based on the number of children in the building i.e. excluding those not pysically present at school
* children attending school that week? Might be better to introduce a lag, as covid infections take 3/5 days to manifest - so possibly better to consider covid infections this week as a proportion of those attending school last?

```{r}
attendance %>% filter(school=="St Joseph's Bingley")
```
```{r}
attendance %>% filter(missing_sessions > 30) %>% group_by(arm) %>% count()
```
```{r}
# remove worst of missing data
attendance <- attendance %>% filter(missing_sessions < 30) 
```
```{r}
attendance <- attendance %>% mutate(pct_attended = sessions_attended / recorded_sessions* 100)
```
```{r}
attendance %>% group_by(hepa_arm) %>% summarise(pct_attended = mean(pct_attended))
```
```{r}
attendance %>% group_by(arm) %>% summarise(pct_attended = mean(pct_attended))
```
```{r}
attended.lm <- lm(pct_attended ~ hepa_arm, data=attendance)
summary(attended.lm)
```
```{r}
attendance <- attendance %>% mutate(pct_illness = illness_absences / recorded_sessions* 100)
```
```{r}
attendance %>% group_by(hepa_arm) %>% summarise(illness_absences = mean(pct_illness))
```
```{r}
attendance %>% group_by(arm) %>% summarise(illness_absences = mean(pct_illness))
```
```{r}
illness.lm <- lm(pct_illness ~ hepa_arm, data=attendance)
summary(illness.lm)
```
```{r}
attendance <- attendance %>% mutate(pct_cov_illness = cov_illness_absences / recorded_sessions* 100)
```
```{r}
attendance %>% group_by(hepa_arm) %>% summarise(illness_absences = mean(pct_cov_illness))
```
```{r}
attendance %>% group_by(arm) %>% summarise(illness_absences = mean(pct_cov_illness))
```
```{r}
cov_illness.lm <- lm(pct_cov_illness ~ hepa_arm, data=attendance)
summary(cov_illness.lm)
```
