---
title: "Bradford School Attendance Data - Problem Schools"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

<!-- Summary: Exploration of issues with the initial attendance data provided by Bradford Council IMT -->

```{r, echo=FALSE, results="hide"}

library(tidyverse)
library(foreach)
library(iterators)
library(gt)

# load attendance data
attendance <- read_csv("data/brad_la_school_attendance_full.csv")

# add study arms to data
attendance$arm <- 0
arm1 = c(
  "Horton Grange Primary School", "Bradford Academy", 
  "Cavendish Primary School", "St Matthew's Catholic Primary School", 
  "Farnham Primary School", "St Joseph's Catholic Primary School (Bingley)", 
  "Green Lane Primary School", "Westbourne Primary School", 
  "Margaret McMillan Primary School", "Beckfoot Nessfield", 
  "Beckfoot Priestthorpe Primary School & Nursery"
)
attendance[attendance$school %in% arm1, "arm"] <- 1
arm2 = c(
  "Eastwood Community School", "Shirley Manor Primary School",  
  "Swain House Primary School", "Bowling Park Primary School", 
  "Burley and Woodhead CofE Primary School", "Crossflatts Primary School", 
  "Copthorne Primary School", "St Columba's Catholic Primary School", 
  "St Joseph's Catholic Primary School (Bradford)", 
  "St William's Catholic Primary School"
)
attendance[attendance$school %in% arm2, "arm"] <- 2
arm3 = c(
  "Riddlesden St Mary's CofE Primary School", "Newby Primary School", 
  "Thackley Primary School", "Addingham Primary School", 
  "The Academy At St. James", "St Paul's CofE Primary School", 
  "Clayton Village Primary School", "Sandal Primary School", 
  "Grove House Primary School", "Saltaire Primary School"
)
attendance[attendance$school %in% arm3, "arm"] <- 3
attendance <- attendance %>% mutate(hepa_arm = arm == 2)

# convert week_start to date format
attendance <- attendance %>% 
  mutate(week_start = as.Date(week_start, format="%m/%d/%y"))

# remove stray october date
attendance <- attendance %>% filter(week_start > "2021-09-01")

# rename St Joseph's schools to make easier to differentiate
attendance <- attendance %>% 
  mutate(school = replace(school, 
                          school == "St Joseph's Catholic Primary School (Bingley)",  
                          "St Joseph's Bingley")) %>%
  mutate(school = replace(school, 
                          school == "St Joseph's Catholic Primary School (Bradford)",  
                          "St Joseph's Bradford"))

# shorten down school names
shorten_name <- function(name) {
    words = unlist(strsplit(name, " "))
    if(length(words) >= 3) {
      paste(words[1:3], collapse=" ")
    } else {
      paste(words, collapse=" ")
    }
}
attendance <- attendance %>% 
  mutate(school = sapply(school, shorten_name))

# add extra aggregate data and reorder
attendance <- attendance %>%
  mutate(other_attendance = rowSums(across(c(L, B, D, J, P, V, W)))) %>%
  mutate(covid_absences = rowSums(across(c(I02, X06, X07, X02, X09, X08, X)))) %>%
  mutate(all_sessions = possible_sessions + covid_absences + I01 + Y + `#`) %>%
  mutate(diff = total_sessions - all_sessions) %>%
  relocate(other_attendance, .after=pm) %>%
  relocate(covid_absences, .after=U) %>%
  relocate(all_sessions, .after=total_sessions) %>%
  relocate(diff, .after=all_sessions)

str(attendance)
```

The following are my notes breaking down the various issues/inconsistencies between the different school's reporting of attendance data. The issues are broken down by individual school with the relevant raw data, but for clarity I've written a quick blurb about the two main issues we're looking at.

## Main issues:

### 1. Missing records:

For some schools' records, when you sum together all the attendance codes "/","\","L", "B", ... through to "#", and compare this figure with the total number of sessions (number of pupils on roll x 14 sessions per week) there are several (sometimes hundreds) of sessions missing. Put another way, there is a 14 character attendance string that should code the attendance info for every student - in several cases these strings are either missing information outright, or include characters that don't correspond to one of the accepted attendance codes.

For some of these examples, we have notes from Bradford Council specifying the issue e.g. the attendance string only details 6 sessions as opposed to the full 14. Others there are so many missing records we can't provide this sort of detail.

### 2. Code "I" vs code "I01":

Before the pandemic, all illness related absences were recorded as a "code I". Following the pandemic, the DfE introduced two new illness codes "I01 - Illness": any non-covid related illness and "I02" - confirmed cases of covid, in an attempt to differentiate covid/non-covid cases. Though unclear, the intention seems to be for the "I01" code to replace the plain "I" code.

Some schools, however, are reporting absences under code "I" and code "I01" each week, which begs the question - what signifies the difference between the absences they are recording under these two codes? Are they just choosing I/I01 at random to signify children with an illness that isn't covid? If not, what difference do they think they're coding?

The schools using both codes seem to correlate strongly with the schools with missing data - so the issues seem to cluster around specific schools. Also, a couple of schools are reporting huge numbers of absences under the I01 code - far more than you would expect to be absent with non covid related illnesses!

Now on to the specific detail for each school:

## Beckfoot Priestthorpe Primary:

### Missing session attendance records for the following weeks:

9/20/21 - 203 missing - no explanation

### Usage of codes I / I01:

Seem to mainly use code I when recording illness absences, but occasional missed sessions recorded under I01. Also a huge spike in covid and non-covid related illness in weeks 04/10 and 11/10 - what's that about?!

```{r, echo=FALSE}
attendance %>% filter(school == "Beckfoot Priestthorpe Primary") %>% 
  select(week_start, I, I01, I02)
```

## Bradford Academy:

### Missing session attendance records for the following weeks:

* 9/6/21 - 6 missing - string for one pupil only 6 sessions long, not 14
* 9/13/21 - 10 missing - string for 1 pupils contains 10 x - (missing marks)
* 9/20/21 - 10 missing - string for 1 pupils contains 10 x - (missing marks)
* 9/27/21 - 10 missing - string for 1 pupils contains 10 x - (missing marks)
* 10/4/21 - 10 missing - string for 1 pupils contains 10 x - (missing marks)
* 10/11/21 - 10 missing - string for 1 pupils contains 10 x - (missing marks)
* 10/18/21 - 164 missing - no specific notes why 
* 11/1/21 - 6 missing - string for one pupil only 6 sessions long, not 14

The 6/10 missing entries aren't so much of an issue - it's predominantly the 164 missing sessions from 18/10/21 that we're concerned about

### Usage of codes I / I01:

Absences recorded under code I AND code I02. Very large numbers of children absent with non-covid related illness code I01 - far higher than the records with I (original illness code) and I02 (confirmed covid absence). Seems unusually high.

```{r, echo=FALSE}
attendance %>% filter(school == "Bradford Academy") %>% 
  select(week_start, I, I01, I02)
```

## Burley and Woodhead:

### Missing attendance records for the following weeks - no notes on reason:

* 9/20/21 - 	11 missing
* 10/18/21 - 	760 missing - Incomplete string for majority of pupils (6 not 14); duplicate string issue

### Usage of codes I / I01:

Seem to mostly use code I01, but some use of code I

```{r, echo=FALSE}
attendance %>% filter(school == "Burley and Woodhead") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```

## Clayton Village Primary:

352 missing sessions for w/c 9-Sep-21 - Brad Council notes say "176 pupils with '--' in string"

## Farnham Priamry School:

### Missing attendance records for the following weeks - no notes on reason:

* 9/6/21 - 3 missing
* 9/13/21 - 8 missing
* 9/20/21 - 24 missing
* 10/4/21 - 28 missing
* 10/11/21 - 1 missing
* 10/18/21 - 16 missing
* 11/1/21 - 27 missing
* 11/8/21 - 37 missing
* 11/15/21 - 4 missing
* 11/22/21 - 150 missing - this is the major concern
* 11/29/21 - 10 missing
* 12/13/21 - 1 missing

### Usage of codes I / I01:

Code I seems to be predomenantly used to code general illness - though there are a couple of examples where both I and I01 are used:

```{r, echo=FALSE}
attendance %>% filter(school == "Farnham Primary School") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```

## Grove House Primary:

### Usage of codes I / I01:

Large numbers of absences recorded under regular I illness code. Very few absences seem to be recorded under I02 (confirmed covid) - more seem to be recorded under I01 (illness not covid). Is I01 being used as the covid illness code?

```{r, echo=FALSE}
attendance %>% filter(school == "Grove House Primary") %>% 
  select(week_start, I, I01, I02)
```

## Margaret McMillan Primary School:

### Missing attendance records for the following weeks - no notes on reason:

* 9/6/21 - 432 missing - this is the major concern
* 9/27/21 -	13 missing
* 10/4/21 -	50 missing
* 10/11/21 - 10 missing
* 10/18/21 - 14 missing
* 11/1/21 -	1 missing
* 11/8/21 -	17 missing

### Usage of codes I / I01:

Code I seems to be predominantly used to code general illness - though there are a few examples where both I and I01 are used. They all look a little unusual too - for the weeks beginning 6/9/21, 20/9/21, 01/11/21 the I01 codes look very similar in number to I02. The 75 sessions missed for I01 w/c 13/09/21 looks very high.

```{r, echo=FALSE}
attendance %>% filter(school == "Margaret McMillan Primary") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```

## Newby Primary School:

### Missing attendance records for the following weeks - no notes on reason:

* 9/6/21 - 	16 missing
* 9/13/21 - 	12 missing
* 9/20/21 - 	6 missing
* 9/27/21 - 	2 missing
* 11/8/21 - 	12 missing
* 11/15/21 - 	2 missing
* 11/22/21 - 	12 missing
* 11/29/21 - 	4 missing
* 12/6/21 - 	2 missing
* 12/13/21 - 	11 missing

(We don't need to be too concerned about these - we can cope with this sort of level of missing data)

### Usage of codes I / I01:

Code I01 seems to be predominantly used, but there are still some absences recorded under code I each week.

```{r, echo=FALSE}
attendance %>% filter(school == "Newby Primary School") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```

## Saltaire Primary School:

### Missing attendance records for the following weeks - no notes on reason:
 
* 9/6/21 - 	24 missing
* 9/13/21 - 	10 missing
* 9/20/21 - 	27 missing
* 9/27/21 - 	26 missing
* 10/18/21 - 	1 missing
* 11/1/21 - 	6 missing
* 12/13/21 - 	280 missing - mainly worried about this!

### Usage of codes I / I01:

Code I01 seems to be predominantly used, but there are still some absences recorded under code I each week.
 
```{r, echo=FALSE}
attendance %>% filter(school == "Saltaire Primary School") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```
 
## Sandal Primary School:
 
### Missing attendance records for the following weeks - no notes on reason:
 
* 9/6/21 - 	9 missing
* 9/13/21 - 	38 missing
* 9/20/21 - 	30 missing
* 10/4/21 - 	20 missing
* 10/11/21 - 	5 missing
* 10/18/21 - 	30 missing
* 11/1/21 - 	30 missing
* 11/8/21 - 	30 missing
* 11/15/21 - 	30 missing
* 11/22/21 - 	20 missing
* 11/29/21 - 	61 missing
* 12/6/21 - 	41 missing
* 12/13/21 - 	61 missing

### Usage of codes I / I01:

No idea what's going on here, I and I01 seem to be used interchangeably.

```{r, echo=FALSE}
attendance %>% filter(school == "Sandal Primary School") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```

## Shirley Manor Primary:
 
### Usage of codes I / I01:

I and I01 seem to be used interchangeably. Also, in latter weeks I02 absences drop to zero, but plenty still recorded under I and I01 - doesn't look right.
 
```{r, echo=FALSE}
attendance %>% filter(school == "Shirley Manor Primary") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```

## St Columba's Catholic:

### Usage of codes I / I01:

I and I01 seem to be used interchangeably. I02 absences are zero almost every week other than w/c 18/10 and 01/11 - seems odd. Is I01 being used to record covid absences?

```{r, echo=FALSE}
attendance %>% filter(school == "St Columba's Catholic") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```

## St Joseph's Bradford

### Usage of codes I / I01:

I and I01 seem to be used interchangeably. I02 absences are zero almost every week other than w/c 15/11, 22/11 and 29/11 - seems odd. Is I01 being used to record covid absences?

```{r, echo=FALSE}
attendance %>% filter(school == "St Joseph's Bradford") %>% 
  select(week_start, I, I01, I02) %>%
  filter(I01 > 0)
```
 
## St Joseph's Bingley
 
### Usage of codes I / I01:
 
Notably different to St Joseph's Bradford! I01 and I02 predominantly seem to be used as expected (non-covid/covid respectively). The majority of weeks, only a small handful of sessions are recorded under code I - however, the final 4 weeks of the year (w/c 22/11, 29/11, 06/12, 13/12) the number of sessions recorded under code I suddenly increases - why? 

```{r, echo=FALSE}
attendance %>% filter(school == "St Joseph's Bingley") %>% 
  select(week_start, I, I01, I02) 
```
## Westbourne Primary School
 
* 115 sessions missing for w/c 13-Sep-21 - no notes why
350 missing sessions for w/c 9-Sep-21 - Brad Council notes say "176 pupils with '--' in string"
