---
title: "exploring_individual_data"
author: "Sam Relins"
date: '2022-07-11'
output: html_document
---

<!-- Summary: Exploration, cleaning and transformation of latest version of attendance data provided by Bradford Council IMT. Cleaning and transformation script extracted to stand-alone R file `clean_and_anonymise_attendance.R` -->

```{r}
library(tidyverse)
library(stringr)
library(foreach)
library(doParallel)
library(gridExtra)
setwd("~/Documents/BIHR/class_act/")
```

```{r}
raw_attendance <- read_csv("data/raw_attendance_sep_2021_jun_2022.csv") %>%
    select(School, WeekStart, WeekMarks)
```

## Fill in missing characters in attendance strings

Some strings < 14 chars long. Fill the missing chars according to the following rules:

    * Some strings are 10 chars long, and so are clearly missing the weekend planned closure chars ("####"). If the string is 10 chars long and doesn't contain any "#"s then add "####" to the end of the string
    * Some strings are 11-13 chars long ending with one or more "#" - assumption is that there are simply a few missing "#" - add the requisite number of "#"s to make the string up to 14
    * Any remaining, add "_"s to represent missing strings that can't be inferred as "#"s 

```{r}
add_missing_att_chars <- function(att_string) {
    if (is.na(att_string)) {
        output <- "______________"
    } else if (nchar(att_string) == 14) {
        output <- att_string
    } else if (nchar(att_string) == 10 & str_count(att_string, "#") == 0) {
        output <- paste(att_string, "####", sep="")
    } else if (nchar(att_string) > 10 & str_sub(att_string, start=-1) == "#") {
        tail_chars <- str_dup("#", 14 - nchar(att_string))
        output <- paste(att_string, tail_chars, sep="")
    } else {
        tail_chars <- str_dup("_", 14 - nchar(att_string))
        output <- paste(att_string, tail_chars, sep="")
    }
    return(output)
}

correct_att_vector <- function(att_vector) {
    output <- foreach(i=1:length(att_vector)) %do% 
        add_missing_att_chars(att_vector[i])
    return(unlist(output))
}

registerDoParallel(8)
raw_attendance <- raw_attendance %>% 
    mutate(att_string = correct_att_vector(WeekMarks))
stopImplicitCluster()
```

## Aggregate records by School/Week

* Sum over columns that count different attendance types
* paste attendance strings into one long string to process

```{r}
attendance <- raw_attendance %>%  
    group_by(School, WeekStart) %>% 
    summarise(pupils = n(),
              att_string = paste(att_string, collapse=""))
```

## Illness columns FYI

Illness column a bit funny - rule seems to be:
    - `*Illness` is all the "=" (I01 - Illness not covid) sessions
    - `Illness Confirmed Coronavirus` is all the ";" (I02 - confirmed covid) 
    - `Illness` is a sum of all the "I"/";"/"=" sessions
    
## Count attendance chars

* count over chars and populate columns headed with absence codes

```{r}
att_chars <- c(
    "_", "-", ";", "\\(", "\\)", "\\[", "\\]", "\\{", "\\}", "\\*", "/", "\\\\",
    "#", "^", "\\+", "=", "B", "C", "D", "E", "G", "H", "I", "J", "L", "M", "N", 
    "O", "P", "R", "S", "T", "U", "V", "W", "X", "Y", "Z"
)

char_descriptions <- c(
    "Unk", "Unk", "I02", "X02", "Unk", "Unk", "X09", "X06", "X07", "Unk", "am", 
    "pm","#", "X08", "X01", "I01", "B", "C", "D", "E", "G", "H", "I", "J", "L", 
    "M", "N", "O", "P", "R", "S", "T", "U", "V", "W", "X", "Y", "Z"
)

unique_descriptions <- unique(unlist(char_descriptions))
for (description in unique_descriptions) {
    attendance[description] <- 0
}

for (i in 1:length(att_chars)) {
    char <- att_chars[i]
    char_count <- str_count(attendance$att_string, char)
    desc <- char_descriptions[i]
    attendance[desc] <- attendance[desc] + char_count 
}
```

## Tidy remaining data

* Shorten School names
* add different study arms
* Format week start column as date

```{r}
attendance <- attendance %>% 
  mutate(WeekStart = as.Date(WeekStart, format="%d/%m/%y"))
drop_schools <- c("Margaret McMillan Primary School", 
                  "Beckfoot Priestthorpe Primary School & Nursery")

attendance <- attendance %>%  
    filter(!(School %in% drop_schools))
  
attendance$arm <- "None" 
arm1 = c(
  "Horton Grange Primary School", "Bradford Academy", 
  "Cavendish Primary School", "St Matthew's Catholic Primary School", 
  "Farnham Primary School", "St Joseph's Catholic Primary School (Bingley)", 
  "Green Lane Primary School", "Westbourne Primary School", 
   "Beckfoot Nessfield"
)
attendance[attendance$School %in% arm1, "arm"] <- "Control" 
arm2 = c(
  "Eastwood Community School", "Shirley Manor Primary School",  
  "Swain House Primary School", "Bowling Park Primary School", 
  "Burley and Woodhead CofE Primary School", "Crossflatts Primary School", 
  "Copthorne Primary School", "St Columba's Catholic Primary School", 
  "St Joseph's Catholic Primary School (Bradford)", 
  "St William's Catholic Primary School"
)
attendance[attendance$School %in% arm2, "arm"] <- "HEPA" 

arm3 <- c(
  "Grove House Primary School", "The Academy At St. James", 
  "Newby Primary School", "Addingham Primary School", 
  "Clayton Village Primary School", "Riddlesden St Mary's CofE Primary School",  
  "Saltaire Primary School")
arm3_activation_dates <- as.Date(c(
    "2021-12-12", "2021-12-12", "2021-12-12", "2021-12-12", "2022-02-13", 
    "2022-03-20", "2022-03-20"
))
arm_3_missing <- c(
  "Thackley Primary School",   "St Paul's CofE Primary School",  
  "Sandal Primary School"
)
for (i in 1:length(arm3)) {
    school_mask = attendance$School == arm3[i]
    date_mask = attendance$WeekStart > arm3_activation_dates[i]
    is_uvc = school_mask & date_mask
    attendance[is_uvc, "arm"] <- "UVC" 
}
attendance[attendance$arm == "None", "arm"] <- "Control" 

attendance <- attendance %>% mutate(HEPA = arm == "HEPA",
                                    UVC = arm == "UVC")
attendance <- attendance %>% 
  mutate(School = replace(School, 
                          School == "St Joseph's Catholic Primary School (Bingley)",  
                          "St Joseph's Bingley")) %>%
  mutate(School = replace(School, 
                          School == "St Joseph's Catholic Primary School (Bradford)",  
                          "St Joseph's Bradford")) %>% 
  mutate(School = replace(School, 
                          School == "The Academy At St. James",  
                          "Academy At St.James"))

shorten_name <- function(name) {
    words = unlist(strsplit(name, " "))
    if(length(words) >= 3) {
      paste(words[1:3], collapse=" ")
    } else {
      paste(words, collapse=" ")
    }
}

attendance <- attendance %>%
  mutate(School = sapply(School, shorten_name))

attendance <- attendance %>% select(-att_string)
```

## Add figures for present in school and illness absence

```{r}
in_school_codes = c("am", "pm", "L", "U")
illness_codes = c("I", "I01", "I02", "X02")
attendance <- attendance %>% 
    mutate(in_school = rowSums(across(in_school_codes)),
           ill = rowSums(across(illness_codes))) %>% 
    mutate(pct_in_school = in_school / (pupils * 14 - `#`),  
           prop_absent_ill = ill /in_school ) %>%
    replace(is.na(.), 0)
```

```{r}
attendance %>% group_by(School) %>% count()
```
```{r}
write_csv(attendance, "~/Documents/BIHR/class_act/data/checkpoints/attendance_ckpt_9_dec_22.csv")
```

Load in anonymisation codes and link with data:

```{r}
school_codes <- read_csv("~/Documents/BIHR/class_act/data/school_keys.csv") %>%
    select(School_AnonID_old, School, msoa)  %>% 
    drop_na() 
anon_attendance <- left_join(attendance, school_codes, by="School") %>%
    rename(closed = `#`, School_AnonID = School_AnonID_old)
```
```{r}
anon_attendance %>% group_by(School_AnonID) %>% count()
```

```{r}
anon_attendance %>% 
    ungroup() %>% 
    relocate(School_AnonID, msoa) %>% 
    select(-School) %>% 
    write_csv("~/Documents/BIHR/class_act/data/checkpoints/anon_attendance_ckpt_9_dec.csv")
```

## School Closures

find any dates where attendance is zero

```{r}
attendance %>% filter(pct_in_school == 0) %>% group_by(WeekStart) %>% count()
```
remove school/weeks for closures:

```{r}
attendance <- attendance %>% filter(pct_in_school > 0)
```

## Errors

plenty of errors - particularly in last week of data - probably due to lack of QA
drop that week

```{r}
attendance <- attendance %>% filter(WeekStart != max(WeekStart)) 
```

```{r}
attendance %>% 
    filter(Unk > 0) %>% 
    mutate(pct_unk = Unk / (pupils *  14) * 100) %>%
    select(WeekStart, School, Unk, pct_unk) %>% 
    arrange(desc(pct_unk)) 
```
for now, bin any entries with > 1% missing data:

```{r}
attendance <- attendance %>% filter((Unk / (pupils *  14)) < 0.01) 
```

## Attendance Patterns

Plot attendance by week

```{r}
(ggplot(data = attendance) 
    + geom_point(mapping=aes(x=WeekStart, y=pct_in_school)))
```

definite break between attendance patterns pre-post 1st Nov 2021

```{r}
(ggplot(data = attendance %>% mutate(pre_nov = WeekStart < as.Date("2021-11-01"))) 
    + geom_boxplot(mapping=aes(x=WeekStart, y=pct_in_school, color=pre_nov))
    + geom_point(mapping=aes(x=WeekStart, y=pct_in_school, color = pre_nov),alpha=0.2))
```
Bradford Academy Seems to account for majority of outlying attendance

```{r}
(ggplot() 
    + geom_smooth(data=attendance %>% 
                     filter(School != "Bradford Academy"), 
                 mapping=aes(x=WeekStart, y=pct_in_school))
    + geom_point(data=attendance,
                 mapping=aes(x=WeekStart, y=pct_in_school),
                 color="blue",
                 alpha=0.2)
    + geom_line(data=attendance %>% 
                     filter(School == "Bradford Academy"), 
                 mapping=aes(x=WeekStart, y=pct_in_school),
                 color="red"))
```
## Illness patterns 

Plot illness by week

```{r}
(ggplot(data = attendance) 
    + geom_point(mapping=aes(x=WeekStart, y=prop_absent_ill))
    + geom_smooth(mapping=aes(x=WeekStart, y=prop_absent_ill)))
```

```{r}
attendance %>% 
           mutate(month_year = as.Date(format(attendance$WeekStart, "01-%m-%Y")))
```
```{r}
attendance %>% 
    mutate(month_year = as.Date(format(attendance$WeekStart, "%Y-%m-01"))) %>%
    select(month_year)
```

```{r}
library(gridExtra)
plot_1 <- (ggplot(
    data=attendance %>% 
        mutate(month_year = as.Date(format(attendance$WeekStart, "%Y-%m-01")))
    ) + geom_boxplot(mapping=aes(x=month_year,  
                                 y=prop_absent_ill,  
                                 color=arm,  
                                 group=interaction(month_year, arm)), 
                     outlier.shape=NA
                   )
    + scale_y_continuous(limits = c(0, 0.16))
 )

plot_2 <- (ggplot(data = attendance)  
           + geom_smooth(mapping=aes(x=WeekStart, y=prop_absent_ill, color=arm)))

grid.arrange(plot_1, plot_2, nrow=2)
```

```{r}
(ggplot(data = attendance ) 
    + geom_point(mapping=aes(x=WeekStart, y=prop_absent_ill, color=arm))
    + geom_smooth(mapping=aes(x=WeekStart, y=prop_absent_ill, color=arm)))
```

```{r, fig.width=3, fig.height=1}
(ggplot(data = attendance) 
    + geom_smooth(mapping=aes(x=WeekStart, y=prop_absent_ill, color=arm)))
```

```{r}
plot_1 <- (ggplot(data=attendance)
    + geom_boxplot(mapping=aes(x=WeekStart,  
                                 y=prop_absent_ill,  
                                 group=WeekStart),  
                   outlier.shape=NA,
                   color="darkblue"
                   )
    + scale_y_continuous(limits = c(0, 0.16))
 )
plot_1
```
```{r}
attendance %>% filter(prop_absent_ill > 0.1) %>%
    group_by(School) %>% count() %>% arrange(desc(n))
```
```{r}
plot_schools = unlist(attendance %>%   
                          filter(prop_absent_ill > 0.1) %>%   
                          group_by(School)  %>%    
                          count() %>%    
                          filter(n > 1) %>%   
                          select(School))
length(plot_schools)
```
```{r}
(ggplot() 
    + geom_line(data = attendance %>% filter(School %in% plot_schools), 
                mapping=aes(x=WeekStart, y=prop_absent_ill, color=School))
    + geom_point(data = attendance %>% filter(!(School %in% plot_schools)), 
                mapping=aes(x=WeekStart, y=prop_absent_ill)))
```
```{r}
(ggplot() 
    + geom_line(data = attendance %>% filter(School == "Bradford Academy"), 
                mapping=aes(x=WeekStart, y=prop_absent_ill, color=School))
    + geom_point(data = attendance %>% filter(!(School == "Bradford Academy")), 
                mapping=aes(x=WeekStart, y=prop_absent_ill)))
```

```{r}
(ggplot() 
    + geom_line(data = attendance %>% filter(School == "Bradford Academy"), 
                mapping=aes(x=WeekStart, y=prop_absent_ill, color=School))
    + geom_point(data = attendance %>% filter(!(School == "Bradford Academy")), 
                mapping=aes(x=WeekStart, y=prop_absent_ill)))
```
```{r}
anon_attendance %>% filter(is.na(School_AnonID))
```

